Constraining Black Hole Growth with High-z Quasar Observations

### 2022-07-28 ms.tex
version 0728 1. title and shortened abstract 2. QLF construction statements in introduction, and small modification in blue to make the logic more fluent.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of black hole mass function (BHMF) and the quasar luminosity function (QLF)
plays an important role in the establishment of BH-galaxy coevolution at high redshifts ($z\gtrsim$ 6).
However, understanding the accretion and radiation properties of growing BHs in the high-$z$ environments 
still remains challenging.
In this work, based on a semi-analytic BH seed formation model, 
we relate BH seeds with the growth tracks of their parent halos. 
We then construct a model of BH mass and luminosity evolution 
characterizing the episodical BH accretion and Eddington ratio distribution function (ERDF).
The model parameters are fitted to reproduce the observed BHMF and QLF of the quasar population at $z\sim$ 6.
Our best-fit models suggest that BHs grow with episodical bursts and quescent phases, with time duration 
typically in units of 20 $\sim$ 30 Myr. 
The duty cycle of active accretion is $f_{\rm duty}\sim 0.15$ for massive BHs with $M\gtrsim 10^8~\Msun$, 
boosted from that in the whole population by nearly a factor of two. 
We find that the intrinsic ERDF of the whole quasar population is well presented by a Schechter shape, 
indicating a large number of low Eddington ratio objects missed by the current flux-limited quasar selections.
We give prediction of the early evolution of BHMF and QLF,
as well as the growth patterns of individual BHs, 
which will likely be unveiled by the James Webb Space Telescope (JWST) and future deep and wide-field surveys.
%  with, e.g. 
% the , Roman Space Telescope (RST), and Euclid.
\end{abstract}

@@

\vspace{5mm}
\section{Introduction} \label{sec:intro}
The growth of massive black holes (BHs) in active galactic nuclei (AGN) is elegantly related to their luminous accretion phases
across the cosmic time \citep{1982MNRAS.200..115S}.
%,1992MNRAS.259..725S}.
Comparing the mass density of massive BHs in the local universe and the mass accreted onto BH inferred from
the integration of the quasar luminosity function based on multi-wavelength observations,
the cosmic BH accretion history and their radiative efficiency are nicely constrained 
\citep[e.g.][]{1971ApJ...170..223C,1992MNRAS.259..725S,2002MNRAS.335..965Y,2004MNRAS.351..169M,2008MNRAS.388.1011M,2004MNRAS.354.1020S,
2009ApJ...690...20S,2014MNRAS.439.2736D,2014ApJ...786..104U,2017A&A...600A..64T}.

Extrapolation of this argument toward higher redshifts beyond $z\sim$ 6 is limited because of the current capability for
high-$z$ quasar observations \citep{2013ApJ...768..105M,2016ApJ...829...33Y,2019ApJ...884...30W,2019BAAS...51c.121F}.
% \red{KI: you could find better reference in this Decadal Survey?}.
Nevertheless, currently the $z\sim 6$ quasar population is well constrained by multiband surveys
\citep{2008AJ....135.1057J,2010AJ....139..906W,2016arXiv161205560C,2016ApJ...833..222J,2018PASJ...70S..35M,2019ApJ...883..183M,2019AJ....157..168D}
along with the quasar luminosity function \citep[e.g.,][]{2018ApJ...869..150M}.
%and the virial BH mass estimates
%from the \ion{Mg}{2} 2798 $\rm{\AA}$ line width 
%where 
Deep spectroscopic observations of high-$z$ quasars enable us to measure the virial BH mass 
with the \ion{Mg}{2} single-epoch method and bring insights of the mass distribution
\citep{2007AJ....134.1150J,2007ApJ...669...32K,2010AJ....140..546W,2018Natur.553..473B,2019ApJ...880...77O,2019ApJ...873...35S}. 
Moreover, in the era of the James Webb Space Telescope (JWST) and forthcoming facilities,
e.g., the Roman Space Telescope (RST) and Euclid,
infrared imaging and spectroscopic observations will unveil a wealth of information on the
high-$z$ quasar properties and their environments 
\citep{2019BAAS...51c..45R, 2019arXiv190205569A, 2011arXiv1110.3193L}. 
Deep surveys of high-$z$ quasars and their host galaxies will shed light on the early BH evolution, 
also help answering the intriguing questions e.g., 
the existence of supermassive BHs in the early universe within one billion year 
\citep{2012Sci...337..544V,2013ASSL..396..293H,2020ARA&A..58...27I}, 
and the establishment of BH-galaxy coevolution 
\citep[e.g.,][]{2013ApJ...773...44W,2017ApJ...851L...8V,2021ApJ...914...36I,2022ApJ...927..237I,2022MNRAS.511.3751H}.
% \red{cite Wang+2013; Venemans+2017, Izumi et al. 2021, Inayoshi+2022 simulation, Habouzit+2022}.

Formation of seed BHs is the starting point to characterize the early assembly of massive BH populations.
Their formation pathway and expected mass range have been extensively investigated by numerical simulations and
semi-analytical studies
%From Pop III star remnants ($10-10^3 ~\Msun$) to direct collapse supermassive stars (10$^4-10^5 ~\Msun$),
%the properties of high-$z$ BH seeds are extensively modeled by simulations and semi-analytical works
%\citep[e.g.,][Toyouchi 2022]{2001ApJ...546..635O,2002Sci...295...93A,2006ApJ...652....6Y,2011MNRAS.416.2748I,
%2012ApJ...756...93H,2013ApJ...778..178H,2014MNRAS.439.3798F,
%2014MNRAS.445L.109I,2014MNRAS.445..544S,2014MNRAS.445..107V,2020MNRAS.499.5960S}.
\citep{2006MNRAS.370..289B,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T}.
% \red{KI: I think here we should cite more general citations: Begelman+2006; Tanaka \& Haiman 2009; Hirano+14, 15,
% Toyouchi+2022, one Natarajan+Volonteri's paper discussing the mass function, Valiante+2018, Sassano+21}
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
%factors which modulate the Pop III star formation have been assessed,
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
%irradiation in Lyman-Werner band which photo-dissociates the major coolant $\hh$
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
% \red{KI: add Chon+16},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
% \red{KI: add Inayoshi, Li \& Haiman 18, Schauer +17},
and
(iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
% \red{KI: add Latif's Nature paper, Mayer+10 or 15}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},% \red{KI: add Becerra+15, Latif+16}, 
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
% \red{cite Hosokawa+13, Woods+19?, Schleicher+13, Toyouchi+22}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
% \red{cite Dijkstra+08, Ahn+09, Inayoshi \& Tanaka 15}, 
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the heavy-seed formation
can be accelerated in the rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and the frequency of halo mergers.
%These factors prevail in some environments at high redshifts and facilitate massive Pop III and even supermassive star formation,
%which contribute to the heavy seed scenario of supermassive BH formation at $z\gtrsim$ 6
%\citep[see a review by][]{2020ARA&A..58...27I}.

\red{Introduction of previous studies on QLFs for high-$z$ populations. mention Yung's paper, Ricarte \& Natarajan paper, recent Finkelstein's paper?
Also, I remember one paper of our former postdoc, Kim.
Also, recent Raffaella's student's paper and "highly-empirical paper" you introduced.
Please add more (I might miss some relevant papers.)}
\blue{
% Haiman \& Loeb 1997.
% High-z
}

To construct the picture of BH evolution, the initial mass function (IMF) of BH seeds is required.
Studies have shown that Pop III stars (or the seed BH population) formed in mini-halos is top-heavy, while capped with $\sim 10^3~\Msun$ 
\citep{2014ApJ...781...60H,2015MNRAS.448..568H}.
Focusing on the progenitor halos of supermassive BHs at $z\sim$ 6, 
our recent semi-analytical work produces an IMF of BH seeds ranging from $10^2-10^5~\Msun$ \citep{2021ApJ...917...60L}. 
The IMF is consistent with the three-dimensional radiation hydrodynamical simulations 
constituting the stellar evolution with large scale mass inflow under realistic halo circumstances (Toyouchi et al. 2022).
In this work, we apply the seeding model to examine the BH growth in common regions in the early universe, 
constrained by the observable QLF and BHMF down to $z\sim$ 6, 
% our aim is to understand the growth patterns of early BHs.

@@

The two parameters characterizing the ERDF are given as
$(\lambda_0,\alpha) = (0.89,0.12)$ and $(0.87,0.20)$ for $\fseed=0.1$ and $0.01$, respectively.
With those values, the fraction of super-Eddington accreting BHs is estimated as $P(\lambda>1)=0.064$ and 0.075, respectively. 
While the probability of rapid accretion in each cycle is similar, multiple accretion episodes enlarge the difference.
Moreover, the best fit value of $\delta$ is $(0.001,0.055)$ for the two $\fseed$ cases,
the higher $\delta$ value encourages the growth of small BHs with $M_\bullet <10^8~\Msun$,
so that the ten-fold difference in $\fseed$ is nearly offset.
%Thus for different $\fseed$ values, a similar number density of BHs which rapidly grow to be SMBHs by $z\simeq 6$ is produced, 
%consistent with a general picture for the early assembly of SMBHs 
%\citep[e.g.,][]{2006NewAR..50..665F,2008AJ....135.1057J,2010AJ....139..906W}.
The characteristic Eddington ratio needs to be close to unity, otherwise the highest luminosity and BH mass
in the model would differ from those in observations.
On the other hand, the power-law index of $\alpha$ in the ERDF tends to increase as the seeding fraction decreases,
indicating that fraction of active BHs with $\lambda \gtrsim 1$ is elevated. 
Thus, with lower seeding fractions, the high-$z$ quasar assembly spends a larger fraction of time in active phases 
and reproduce the bulk properties of the BHMF and QLF.

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of black hole mass function (BHMF) and the quasar luminosity function (QLF)
plays an important role in the establishment of BH-galaxy coevolution at high redshifts ($z\gtrsim$ 6).
However, understanding the accretion and radiation properties of growing BHs in the high-$z$ environments 
still remains challenging.
In this work, based on a semi-analytic BH seed formation model, 
we relate BH seeds with the growth tracks of their parent halos. 
We then construct a model of BH mass and luminosity evolution 
characterizing the episodical BH accretion and Eddington ratio distribution function (ERDF).
The model parameters are fitted to reproduce the observed BHMF and QLF of the quasar population at $z\sim$ 6.
Our best-fit models suggest that BHs grow with episodical bursts and quescent phases, with time duration 
typically in units of 20 $\sim$ 30 Myr. 
The duty cycle of active accretion is $f_{\rm duty}\sim 0.15$ for massive BHs with $M\gtrsim 10^8~\Msun$, 
boosted from that in the whole population by nearly a factor of two. 
We find that the intrinsic ERDF of the whole quasar population is well presented by a Schechter shape, 
indicating a large number of low Eddington ratio objects missed by the current flux-limited quasar selections.
We predict the early evolution of BHMF and QLF, as well as the growth patterns of individual BHs, 
which will likely be unveiled by the James Webb Space Telescope (JWST) and future deep and wide-field surveys.
%  with, e.g. 
% the , Roman Space Telescope (RST), and Euclid.
\end{abstract}

@@

\vspace{5mm}
\section{Introduction} \label{sec:intro}
The growth of massive black holes (BHs) in active galactic nuclei (AGN) is elegantly related to their luminous accretion phases
across the cosmic time \citep{1982MNRAS.200..115S}.
%,1992MNRAS.259..725S}.
Comparing the mass density of massive BHs in the local universe and the mass accreted onto BH inferred from
the integration of the quasar luminosity function based on multi-wavelength observations,
the cosmic BH accretion history and their radiative efficiency are nicely constrained 
\citep[e.g.][]{1971ApJ...170..223C,1992MNRAS.259..725S,2002MNRAS.335..965Y,2004MNRAS.351..169M,2008MNRAS.388.1011M,2004MNRAS.354.1020S,
2009ApJ...690...20S,2014MNRAS.439.2736D,2014ApJ...786..104U,2017A&A...600A..64T,2022arXiv220607357S}.

Extrapolation of this argument toward higher redshifts beyond $z\sim$ 6 is limited because of the current capability for
high-$z$ quasar observations \citep{2013ApJ...768..105M,2016ApJ...829...33Y,2019ApJ...884...30W,2019BAAS...51c.121F}.
% \red{KI: you could find better reference in this Decadal Survey?}.
Nevertheless, currently the $z\sim 6$ quasar population is well constrained by multiband surveys
\citep{2008AJ....135.1057J,2010AJ....139..906W,2016arXiv161205560C,2016ApJ...833..222J,2018PASJ...70S..35M,2019ApJ...883..183M,2019AJ....157..168D}
along with the quasar luminosity function \citep[e.g.,][]{2018ApJ...869..150M}.
%and the virial BH mass estimates
%from the \ion{Mg}{2} 2798 $\rm{\AA}$ line width 
%where 
Deep spectroscopic observations of high-$z$ quasars enable us to measure the virial BH mass 
with the \ion{Mg}{2} single-epoch method and bring insights of the mass distribution
\citep{2007AJ....134.1150J,2007ApJ...669...32K,2010AJ....140..546W,2018Natur.553..473B,2019ApJ...880...77O,2019ApJ...873...35S}. 
Moreover, in the era of the James Webb Space Telescope (JWST) and forthcoming facilities,
e.g., the Roman Space Telescope (RST) and Euclid,
infrared imaging and spectroscopic observations will unveil a wealth of information on the
high-$z$ quasar properties and their environments 
\citep{2019BAAS...51c..45R, 2019arXiv190205569A, 2011arXiv1110.3193L}. 
Deep surveys of high-$z$ quasars and their host galaxies will shed light on the early BH evolution, 
also help answering the intriguing questions e.g., 
the existence of supermassive BHs in the early universe within one billion year 
\citep{2012Sci...337..544V,2013ASSL..396..293H,2020ARA&A..58...27I}, 
and the establishment of BH-galaxy coevolution 
\citep[e.g.,][]{2013ApJ...773...44W,2017ApJ...851L...8V,2021ApJ...914...36I,2022ApJ...927..237I,2022MNRAS.511.3751H}.

\blue{
Nevertheless, starting from empirical BH seeding, combinded with growth chanels e.g., merger and accretion modes,
models can in return give prediction of the QLF at high redshifts 
\citep[e.g.,][]{2018MNRAS.474.1995R,2018MNRAS.481.3278R,2021MNRAS.508.2706Y,2022MNRAS.511..616T}. 
The classical way of regulating BH growth is by keeping the tight relation between $M_\bullet$ 
and the parent halo mass or the mass of the stellar bulge
\citep{2007ApJ...669...45H,2013ARA&A..51..511K},
then following the hierarchical DM halo growth
\citep[e.g.,][]{1998ApJ...503..505H,2010ApJ...718..231S,2021ApJ...910L..11K}.
The model parameters can thus be calibrated with low-$z$ BHMF and QLF,
and subsequently predict the higher redshift and faint quassar observations.
On the other hand, studies also conduct extrafpolation on the low-$z$ QLF evolution and forecast the high-$z$ QLF shapes,
which could be investigated by furture surveys
\citep[e.g.,][]{2020MNRAS.495.3252S,2022arXiv220702233F}.
}

\red{Introduction of previous studies on QLFs for high-$z$ populations. mention Yung's paper, Ricarte \& Natarajan paper, recent Finkelstein's paper?
Also, I remember one paper of our former postdoc, Kim.
Also, recent Raffaella's student's paper and "highly-empirical paper" you introduced.
Please add more (I might miss some relevant papers.)}

\blue{
As a starting point to characterize the early assembly of massive BH populations,
the seed BH formation has been extensively investigated by numerical simulations and
semi-analytical studies
}
% Formation of seed BHs is the starting point to characterize the early assembly of massive BH populations.
% Their formation pathway and expected mass range have been extensively investigated by numerical simulations and
% semi-analytical studies
%
%From Pop III star remnants ($10-10^3 ~\Msun$) to direct collapse supermassive stars (10$^4-10^5 ~\Msun$),
%the properties of high-$z$ BH seeds are extensively modeled by simulations and semi-analytical works
%\citep[e.g.,][Toyouchi 2022]{2001ApJ...546..635O,2002Sci...295...93A,2006ApJ...652....6Y,2011MNRAS.416.2748I,
%2012ApJ...756...93H,2013ApJ...778..178H,2014MNRAS.439.3798F,
%2014MNRAS.445L.109I,2014MNRAS.445..544S,2014MNRAS.445..107V,2020MNRAS.499.5960S}.
\citep{2006MNRAS.370..289B,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T}.
% \red{KI: I think here we should cite more general citations: Begelman+2006; Tanaka \& Haiman 2009; Hirano+14, 15,
% Toyouchi+2022, one Natarajan+Volonteri's paper discussing the mass function, Valiante+2018, Sassano+21}
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
%factors which modulate the Pop III star formation have been assessed,
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
%irradiation in Lyman-Werner band which photo-dissociates the major coolant $\hh$
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
% \red{KI: add Chon+16},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
% \red{KI: add Inayoshi, Li \& Haiman 18, Schauer +17},
and
(iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
% \red{KI: add Latif's Nature paper, Mayer+10 or 15}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},% \red{KI: add Becerra+15, Latif+16}, 
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
% \red{cite Hosokawa+13, Woods+19?, Schleicher+13, Toyouchi+22}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
% \red{cite Dijkstra+08, Ahn+09, Inayoshi \& Tanaka 15}, 
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the heavy-seed formation
can be accelerated in the rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and the frequency of halo mergers.
%These factors prevail in some environments at high redshifts and facilitate massive Pop III and even supermassive star formation,
%which contribute to the heavy seed scenario of supermassive BH formation at $z\gtrsim$ 6
%\citep[see a review by][]{2020ARA&A..58...27I}.

\blue{
More explicitly, the initial mass function (IMF) of BH seeds is required to construct the picture of BH evolution.
}
% To construct the picture of BH evolution, the initial mass function (IMF) of BH seeds is required.
Studies have shown that Pop III stars (or the seed BH population) formed in mini-halos is top-heavy, while capped with $\sim 10^3~\Msun$ 
\citep{2014ApJ...781...60H,2015MNRAS.448..568H}.
Focusing on the progenitor halos of supermassive BHs at $z\sim$ 6, 
our recent semi-analytical work produces an IMF of BH seeds ranging from $10^2-10^5~\Msun$ \citep{2021ApJ...917...60L}. 
The IMF is consistent with the three-dimensional radiation hydrodynamical simulations 
constituting the stellar evolution with large scale mass inflow under realistic halo circumstances (Toyouchi et al. 2022).
In this work, we apply the seeding model to examine the BH growth in common regions in the early universe, 
constrained by the observable QLF and BHMF down to $z\sim$ 6, 
% our aim is to understand the growth patterns of early BHs.

@@

The two parameters characterizing the ERDF are given as
$(\lambda_0,\alpha) = (0.89,0.12)$ and $(0.87,0.20)$ for $\fseed=0.1$ and $0.01$, respectively.
With those values, the fraction of super-Eddington accreting BHs is estimated as $P(\lambda>1)=0.064$ and 0.075, respectively. 
While the probability of rapid accretion in each cycle is similar, multiple accretion episodes enlarge the difference.
Moreover, the best fit values of $\delta$ are $(0.001,0.055)$ for the two $\fseed$ cases,
the higher $\delta$ value encourages the growth of small BHs with $M_\bullet <10^8~\Msun$,
so that the ten-fold difference in $\fseed$ is nearly offset.
%Thus for different $\fseed$ values, a similar number density of BHs which rapidly grow to be SMBHs by $z\simeq 6$ is produced, 
%consistent with a general picture for the early assembly of SMBHs 
%\citep[e.g.,][]{2006NewAR..50..665F,2008AJ....135.1057J,2010AJ....139..906W}.
The characteristic Eddington ratio needs to be close to unity, otherwise the highest luminosity and BH mass
in the model would differ from those in observations.
On the other hand, the power-law index of $\alpha$ in the ERDF tends to increase as the seeding fraction decreases,
indicating that fraction of active BHs with $\lambda \gtrsim 1$ is elevated. 
Thus, with lower seeding fractions, the high-$z$ quasar assembly spends a larger fraction of time in active phases 
and reproduce the bulk properties of the BHMF and QLF.
